Baseline Drift Estimation for Air Quality Data Using Quantile Trend Filtering

### 2019-04-18 Manuscript/detrendify_aoas.tex
fixed primal residual
edited timing section

@@

\begin{eqnarray*}
	r_{\text{primal}}^{m} & = & \sqrt{\sum_{w=1}^W\lVert\Mn{\Theta}{w}_{m} - \M{\Theta}_m\rVert_{\text{F}}^2},
\end{eqnarray*}
represents the difference between the trend values in the windows and the consensus trend value. The dual residual at the $m$th iteration,

@@

\Alg{admm} has the following convergence guarantees.

\begin{proposition}
	\label{prop:convergence}
	Let $\{\{\Mn{\Theta}{w}_m\}, \M{\Theta}_m\}$ denote the $m$th collection of iterates generated by \Alg{admm}. Then (i)
	$\lVert \Mn{\Theta}{w}_m - \Mn{U}{w}\M{\Theta}_m \rVert_{\text{F}} \rightarrow 0$ and (ii) $p_m \rightarrow p^\star$, where $p^\star$ is the optimal objective function value of
	\Eqn{quantile_windows} and $p_m$ is the objective function value of \Eqn{quantile_windows} evaluated at $\{\{\Mn{\Theta}{w}_m\}, \M{\Theta}_m\}$.
\end{proposition}

The proof of \Prop{convergence} is a straightforward application of the convergence result presented in Section 3.2 of \cite{boyd2011distributed}.

The quantile trend filtering problem for a single window is a linear program with $N\times J$ parameters (number of observations by number of quantiles), which can be solved in computational time proportional to $(NJ)^3$. By dividing the problem into four smaller problems we can theoretically solve each of the 4 sub-problems 4 times and still reduce the computation time required by a factor of 4. Consequently, solving a large problem using \Alg{admm} should require less computational time than simply using a linear program solver given small enough windows and convergence in few enough iterations. We demonstrate the advantages of \Alg{admm} empirically using timing experiments. For each data size, $n$, 25 datasets were simulated using the peaks simulation design described below. Trends for three quantiles were fit simultaneously: 0.05, 0.1, and 0.15 using $\lambda = n/5$. We use from one to four windows for each data size with an overlap of 500. The windows algorithm was run until the stopping criteria were met using $\epsilon_{abs} = 0.01$ and $\epsilon_{rel} = 0.001$. \Fig{timing} shows that using 4 windows instead of one on data sizes of 55000 provides a factor of 3 decrease in computation time. The timing experiments were conducted on an Intel Xeon based Linux cluster using two processor cores.

>>>>

\Alg{admm} has the following convergence guarantees.
\begin{proposition}
	\label{prop:convergence}
	Let $\{\{\Mn{\Theta}{w}_m\}, \M{\Theta}_m\}$ denote the $m$th collection of iterates generated by \Alg{admm}. Then (i)
	$\lVert \Mn{\Theta}{w}_m - \Mn{U}{w}\M{\Theta}_m \rVert_{\text{F}} \rightarrow 0$ and (ii) $p_m \rightarrow p^\star$, where $p^\star$ is the optimal objective function value of
	\Eqn{quantile_windows} and $p_m$ is the objective function value of \Eqn{quantile_windows} evaluated at $\{\{\Mn{\Theta}{w}_m\}, \M{\Theta}_m\}$.
\end{proposition}
The proof of \Prop{convergence} is a straightforward application of the convergence result presented in Section 3.2 of \cite{boyd2011distributed}.

@@

\begin{eqnarray*}
	r_{\text{primal}}^{m} & = & \sqrt{\sum_{w=1}^W\lVert\Mn{\Theta}{w}_{m} - \Mn{U}{w}\M{\Theta}_m\rVert_{\text{F}}^2},
\end{eqnarray*}
represents the difference between the trend values in the windows and the consensus trend value. The dual residual at the $m$th iteration,

@@

The quantile trend filtering problem for a single window is a linear program with $N\times J$ parameters (number of observations by number of quantiles), which can be solved in computational time proportional to $(NJ)^3$.  Consequently, solving a large problem using \Alg{admm} should require less computational time than solving \Eqn{noncross_trend}, even if the sub-problems are solved sequentially. We demonstrate the advantages of \Alg{admm} through timing experiments (\Fig{timing}). For each data size, 25 datasets were simulated using the peaks simulation design described below. Trends for the fifth, tenth, and fifteenth quantiles were fit simultaneously: using $\lambda = n/5$. We use from one to four windows for each data size with an overlap of 500. The windows algorithm was run until the stopping criteria were met using $\epsilon_{abs} = 0.01$ and $\epsilon_{rel} = 0.001$. \Fig{timing} shows that using 4 windows instead of one on data sizes of 55,000 provides a factor of 3 decrease in computation time. The timing experiments were conducted on an Intel Xeon based Linux cluster using two processor cores.
